Analysis

### 2022-06-04 manuscript/manuscript.Rmd
--max-stored-scores 1000000

CTCF binding sites were detected using the FIMO tool from the MEME v.4.11.2 suite [@Bailey:2009aa] using default parameters. Softmasked genome sequences were downloaded from the USCS Genome Browser database with the exception of T2T genome (chm13v2.0) downloaded from the https://github.com/marbl/CHM13 repository. Motifs were scanned separately for each chromosome (autosomal and sex chromosomes only), and the results were combined. We used human- and mouse-specific matrices for the corresponding genomes (HOCOMOCO, CIS-BP). As JASPAR, CTCFBSDB, and CIS-BP databases provide several CTCF motifs, all database-specific motifs were used to scan the genomes; consequently, the detected CTCF sites may overlap due to motif similarity. We provide data as GenomicRanges objects and BED files with "seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence" columns.

@@

We used the FIMO tool from the MEME suite [@Bailey:2009aa] as one of the oldest and most well-known tool for motif scanning. The tool uses a dynamic programming algorithm to convert log-odds scores of motif matching into p-values, assuming a zero-order background model [@Cuellar-Partida:2012wj]. The default p-value cutoff used to generate the package's data is 1E-4. However, we observed that up to 98% of motifs have detection p-value more than 1E-6 (`Figure 2A`), suggesting the majority of CTCF motifs have relatively poor match. Filtering results at 1E-6 allows to focus on more reliably detected CTCF sites. Furthermore, given some databases provide multiple CTCF PWMs, FIMO detects many overlapping CTCF sites (`Table S1`). Reducing them (merging overlapping CTCF sites) in combination with 1E-6 cutoff filtering reduces the number of CTCF sites making them comparable with previously reported 20,000-50,000 number of CTCF binding sites. Although we provide FIMO results as-is, we recommend filtering and reducing them.

>>>>

CTCF binding sites were detected using the FIMO tool from the MEME v.4.11.2 suite [@Bailey:2009aa] using default parameters and the `--max-stored-scores 1000000` setting to avoid automatic drop of less significant motifs in case of memory limits. Softmasked genome sequences were downloaded from the USCS Genome Browser database with the exception of T2T genome (chm13v2.0) downloaded from the https://github.com/marbl/CHM13 repository. Motifs were scanned separately for each chromosome (autosomal and sex chromosomes only), and the results were combined. We used human- and mouse-specific matrices for the corresponding genomes (HOCOMOCO, CIS-BP). As JASPAR, CTCFBSDB, and CIS-BP databases provide several CTCF motifs, all database-specific motifs were used to scan the genomes; consequently, the detected CTCF sites may overlap due to motif similarity. We provide data as GenomicRanges objects and BED files with "seqnames", "start", "end", "name", "score", "strand", "width", "pvalue", "qvalue", "sequence" columns.

@@

We used the FIMO tool from the MEME suite [@Bailey:2009aa] as one of the oldest and most well-known tool for motif scanning. The tool uses a dynamic programming algorithm to convert log-odds scores of motif matching into p-values, assuming a zero-order background model [@Cuellar-Partida:2012wj]. The default p-value cutoff used to generate the package's data is 1E-4. However, we observed that up to 98% of motifs have detection p-value more than 1E-6 (`Figure 2A`), suggesting the majority of CTCF motifs have relatively poor match. Filtering results at 1E-6 allows to focus on more reliably detected CTCF sites. Furthermore, given some databases provide multiple CTCF PWMs, FIMO detects many overlapping CTCF sites (`Table S1`). Reducing them (merging overlapping CTCF sites) in combination with 1E-6 cutoff filtering yields the number of CTCF sites comparable with previously reported 20,000-50,000 sites. Although we provide FIMO results as-is, we recommend filtering and reducing them.
